CMSAC_reveal_js

### 2025-10-10 CMSAC_reveal_js.qmd
new intro slides

##

@@

@@

## Weighted Kendall
$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$

@@

## Weighted Kendall Example
- If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then:
$$ W = \begin{pmatrix}
3 & 3 & 3 & 3 \\
3 & 1 & 1 & 1 \\
3 & 1 & 1 & 1 \\
3 & 1 & 1 & 0
\end{pmatrix} $$

@@

## Normalizing Mutual Information
- Mutual Information can be normalized by dividing by the joint entropy

@@

- So, the final formula for the metric is:
$$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
## Results

<!-- Should these plots be on separate slides? -->
```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
knitr::include_graphics("tournament_comparison_plot_close.png")
```

## Limitations
- The "known" strengths are assumptions that can be estimated but never truly known
- Ties are not addressed in any of the games
- Ties in the $\hat{r}$ are randomly assigned

## Future Work
- Evaluate structures using various distributions and "known" weights
- Estimate the $\theta$s in a real-world example and compare recent changes to the old format

## References
<!-- I don't know if the references.bib would work here or not -->
- Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.
- Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules.
1st Ed. Cham, Switzerland: Palgrave Macmillan.
- Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an
Operational Research Perspective.” European Journal of Operational Research 324 (1):
1–21.
- Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47
(3-4): 253–62.

## References
- Guiasu, Silviu. 1977. Information Theory with Applications. New York, NY: McGraw-Hill.
- Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan
Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.”
Int. J. Emp. Econ. 01 (02).
- Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking
Teams.” Stat. Modelling 18 (5-6): 411–35.
- Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.”
Comput. Oper. Res. 144 (105821): 105821.

>>>>

@@

## Extreme Tournament

- 4 team single elimination tournament:
  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
- If we only care about the top team winning, seeding structure does not matter
- If we care about who gets second and third, seeding becomes important
  - Would not pair up Work Group and Dream Team in the first round

## Comparing Structures

- Comparing different structures like round robin and single eliminations, which is best?

- Efficacy: Final rankings should match the true rankings of the strength of the competitors

- Effectivity: Interested in correctly ranking the best competitor as first

CITEE

## Efficacy and Effectivity

- Other competitions where the correct ranking of some, but not all, of the competitors is of import
  - Olympic competition ranks the top 3 competitors

- "Salient ranks": the ranks of a competition structure that matter
  - Efficacy: 1 through $n$ matter
  - Effectivity: 1 is only salient rank

- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

@@

## Weighted Kendall

$$ \tau_w = \frac{\sum_{i\neq j} W_{ij} \times \text{sgn}(i-j)\times \text{sgn}(R_i - R_j)}{\sum_{i,j} W_{ij} - \sum_i W_{ii}}$$

@@

## Weighted Kendall Example

-   If $\hat{r} = (2,1,3,4)$ and m = $(3,1,1,0)$, then: $$ W = \begin{pmatrix}
    3 & 3 & 3 & 3 \\
    3 & 1 & 1 & 1 \\
    3 & 1 & 1 & 1 \\
    3 & 1 & 1 & 0
    \end{pmatrix} $$

@@

## Normalizing Mutual Information

-   Mutual Information can be normalized by dividing by the joint entropy

@@

-   So, the final formula for the metric is: $$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$ \## Results

<!-- Should these plots be on separate slides? -->

```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
knitr::include_graphics("tournament_comparison_plot_close.png")
```

## Limitations

-   The "known" strengths are assumptions that can be estimated but never truly known
-   Ties are not addressed in any of the games
-   Ties in the $\hat{r}$ are randomly assigned

## Future Work

-   Evaluate structures using various distributions and "known" weights
-   Estimate the $\theta$s in a real-world example and compare recent changes to the old format

## References

<!-- I don't know if the references.bib would work here or not -->

-   Appleton, David R. 1995. “May the Best Man Win?” Statistician 44 (4): 529.
-   Csato, L. 2021. Tournament Design: How Operations Research Can Improve Sports Rules. 1st Ed. Cham, Switzerland: Palgrave Macmillan.
-   Devriesere, Csató, K., and D. Goossens. 2025. “Tournament Design: A Review from an Operational Research Perspective.” European Journal of Operational Research 324 (1): 1–21.
-   Glenn, W A. 1960. “A Comparison of the Effectiveness of Tournaments.” Biometrika 47 (3-4): 253–62.

## References

-   Guiasu, Silviu. 1977. Information Theory with Applications. New York, NY: McGraw-Hill.
-   Johnson, Sidney, and Rodney Fort. 2022. “Match Outcome Uncertainty and Sports Fan Demand: An Agnostic Review and the Standard Economic Theory of Ports Leagues.” Int. J. Emp. Econ. 01 (02).
-   Lasek, Jan, and Marek Gagolewski. 2018. “The Efficacy of League Formats in Ranking Teams.” Stat. Modelling 18 (5-6): 411–35.
-   Sziklai, Balázs R, Péter Biró, and László Csató. 2022. “The Efficacy of Tournament Designs.” Comput. Oper. Res. 144 (105821): 105821.
